Automatic Amortized Analysis for Imperative Code

### 2014-03-10 main.tex
more cases

\\ \Rule{While}
{ \Gamma, 0 < x ; Q - M_w^2 \vdash S \dashv Q; \Gamma }
{ \Gamma; Q + M_w^1 \vdash \code{while}(0 < x)~S \dashv Q - M_w^3; \Gamma, x \le 0 }

@@

This big step semantics counts the number of iterations of
while loops in a program as it executes it.  Non-deterministically,
the {\sc S:Abort} rule can be called to interrupt a computation.
It has the same effect as raising an exception.  We symbolize a
frozen computation with the notation $\circ$.  In the rules, $X$
stands for either $\circ$ or $H$.

@@

\item Case {\sc S:Seq}.
  By transitivity and induction hypothesis, it is
  easy to get $H'' \models \Gamma''$.  Now we prove
  the quantitative part of the theorem.  We have
  $(q, q') = M_s^1 \cdot (q_1, q'_1) \cdot M_s^2 \cdot (q_2, q'_2)$.
  We assume $\forall i \in \{1, 2\},~M_s^i \ge 0$.
  \begin{itemize}[topsep=0pt]
  \item
    The first case is when $q'_1 > M_s^2 + q_2$ then $q = M_s^1 + q_1$
    and since $q_1 \le \Phi_H(Q - M_s^1)$ (by induction) we have
    $q \le \Phi_H(Q)$.
  \item
    The second case is when $q'_1 \le M_s^2 + q_2$ then $q = M_s^1 + q_1 + M_s^2 + q_2 - q'_1$. \\
    Since $\Phi_H(Q - M_s^1) - \Phi_{H'}(Q') \ge q_1 - q'_1$
    and $q_2 \le \Phi_{H'}(Q' - M_s^2)$ we have:
    \begin{align*}
    q &\le M_s^1 + \Phi_H(Q - M_s^1) - \Phi_{H'}(Q') + M_s^2 + q_2 \\
      &= \Phi_H(Q) + (q_2 - \Phi_{H'}(Q' - M_s^2)) \le \Phi_H(Q).
   \end{align*}
  \end{itemize}
  We obtain $\Phi_H(Q) - \Phi_{H''}(Q'') \ge q - q'$ by simply chaining
  the two induction hypothesis and using the fact that $q - q' =
  (q_1 + M_s^1 - q'_1) + (q_2 + M_s^2 - q'_2)$.

\item Case {\sc S:WhileStop}.
\item Case {\sc S:WhileStep}.

>>>>

\\ \Rule{While}
{ \Gamma, 0 < x ; Q - M_w^1 \vdash S \dashv M_w^2 + Q; \Gamma }
{ \Gamma; Q \vdash \code{while}(0 < x)~S \dashv Q - M_w^3; \Gamma, x \le 0 }

@@

This big step semantics is parameterized by a metric that sets
the cost of each program operation.  Non-deterministically,
the rule {\sc S:Abort} can be called to interrupt a computation.
It has the same effect as raising an exception.  We symbolize a
frozen computation with the notation $\circ$.  In the rules, $X$
stands for either $\circ$ or $H$.

@@

\item Case {\sc S:Seq}.
  By transitivity and induction hypothesis, it is
  easy to get $H'' \models \Gamma''$.  Now we prove
  the quantitative part of the theorem.  We have
  $(q, q') = M_s^1 \cdot (q_1, q'_1) \cdot M_s^2 \cdot (q_2, q'_2)$.
  We assume $\forall i \in \{1, 2\},~M_s^i \ge 0$, the other cases
  are similar.
  \begin{itemize}[topsep=0pt]
  \item
    The first case is when $q'_1 > M_s^2 + q_2$ then $q = M_s^1 + q_1$
    and since $q_1 \le \Phi_H(Q - M_s^1)$ (by induction) we have
    $q \le \Phi_H(Q)$.
  \item
    The second case is when $q'_1 \le M_s^2 + q_2$ then $q = M_s^1 + q_1 + M_s^2 + q_2 - q'_1$. \\
    Since $\Phi_H(Q - M_s^1) - \Phi_{H'}(Q') \ge q_1 - q'_1$
    and $q_2 \le \Phi_{H'}(Q' - M_s^2)$ we have:
    \begin{align*}
    q &\le M_s^1 + \Phi_H(Q - M_s^1) - \Phi_{H'}(Q') + M_s^2 + q_2 \\
      &= \Phi_H(Q) + (q_2 - \Phi_{H'}(Q' - M_s^2)) \le \Phi_H(Q).
   \end{align*}
  \end{itemize}
  We obtain $\Phi_H(Q) - \Phi_{H''}(Q'') \ge q - q'$ by simply chaining
  the two induction hypothesis and using the fact that $q - q' =
  (q_1 + M_s^1 - q'_1) + (q_2 + M_s^2 - q'_2)$.

\item Case {\sc S:WhileStop}.
  This proof goes just like in the {\sc S:Skip} case.

\item Case {\sc S:WhileStep}.
  Here again we assume $\forall i \in \{1, 2, 3\},~M_w^i \ge 0$.
  \begin{itemize}[topsep=0pt]
  \item
    If $q'_1 > M_w^2 + q_2$ then $q = M_w^1 + q_1$ and by induction
    we get $q_1 \le \Phi_H(Q - M_w^1)$, so $q \le \Phi_H(Q)$.
  \item
    If $q'_1 \le M_w^2 + q_2$ then $q = M_w^1 + q_1 + M_w^2 + q_2 - q'_1$.
    And we proceed like in the sequence case.
  \end{itemize}
  We obtain $\Phi_H(Q) - \Phi_{H''}(Q - M_w^3) \ge q - q'$ using the identity
  $q - q' = (M_w^1 + M_w^2 + q_1 - q'_1) + q_2 - q'_2$ and the two
  induction hypothesis
  \begin{itemize}[topsep=0pt]
  \item $\Phi_H(Q - M_w^1) - \Phi_{H'}(Q + M_w^2) \ge q_1 - q'_1$ and
  \item $\Phi_{H'}(Q) - \Phi_{H''}(Q - M_w^3) \ge q_2 - q'_2.$
  \end{itemize}
